---
title: 'Reproducible Research: Project 1'
author: 'Renato Oliveira Fraga'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r}
library(ggplot2)
library(plyr)
```

## Loading and preprocessing the data

```{r}
unzip("repdata_data_activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
SPD <- ddply(data, c("date"),summarise,TS = sum(steps,na.rm=TRUE))
```                   

```{r ,warning=FALSE}

ggplot(data,aes(x=steps, col=date)) + geom_freqpoly(bins=10) + xlab("Total number of steps") + ylab("Frequency")+ ggtitle("Total Steps per Day")

```

Calculate and report the mean and median of the total number of steps taken per day:

- Mean total number of steps taken per day = `r mean(SPD$TS)` 
- Median number of steps taken per day = `r median(SPD$TS)`


## What is the average daily activity pattern?

```{r}
SP5m <- ddply(data, c("interval"),summarise, MS = mean(steps,na.rm=TRUE))
 
ggplot(SP5m,aes(x=interval,y=MS)) + geom_line() + ggtitle("Average steps for each 5-min interval") + xlab("Interval") + ylab("Mean steps")
               
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

- Interval = `r SP5m[which(SP5m$MS==max(SP5m$MS)), "interval"]` 
- Mean = `r SP5m[which(SP5m$MS==max(SP5m$MS)), "MS"]` steps.  

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs):

The dataset has `r nrow(data)-sum(complete.cases(data))` missing values in the dataset.
```{r ,warning=FALSE}
fillData <- aggregate(steps~interval, data=data, FUN=mean)
fillNA <- numeric()
for(i in 1:nrow(data)){
   x <- data[i, ]
   if (is.na(x$steps)){
       steps <- subset(fillData, interval==x$interval)$steps
   } else { 
       steps <- x$steps
   }
   
   fillNA <- c(fillNA, steps)
}
data$steps <- fillNA
ggplot(data,aes(x=steps, col=date)) + geom_freqpoly(bins=10) + xlab("Total number of steps") + ylab("Frequency")+ ggtitle("Total Steps per Day")

```

```{r}
SPD2 <- ddply(data, c("date"),summarise,TS2 = sum(steps,na.rm=TRUE))
```   

Calculate and report the mean and median of the total number of steps taken per day:

- Mean total number of steps taken per day = `r mean(SPD2$TS2)` 
- Median number of steps taken per day = `r median(SPD2$TS2)`


## Are there differences in activity patterns between weekdays and weekends?

```{R}
library(lattice)
data2 <- read.csv("activity.csv")
data2$date<- as.Date(data2$date,"%Y-%m-%d")
day <- weekdays(data2$date)

typeDay <- vector()
for (j in 1:nrow(data2)) {
  if (day[i] == "Saturday") {
    typeDay[j] <- "Weekend"
  } else if (day[j] == "Sunday") {
    typeDay[j] <- "Weekend"
  } else {
    typeDay[j] <- "Weekday"
  } 
  
}
data2$typeDay <- typeDay
data2$typeDay <- factor(data2$typeDay)
stepsByDay <- aggregate(steps ~ interval + typeDay, data = data2, mean)
names(stepsByDay) <- c("interval", "typeDay", "steps")
xyplot(steps ~ interval | typeDay, stepsByDay, type = "l",
       layout = c(1, 2),
       xlab = "Interval", ylab = "Number of steps")
```

According to the results, the activity is higher on the weekends.






